Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-25 docs/manuscript/manuscript.tex
manu updates

\section{Introduction}
%%(Roughly 3 paragraphs)
In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking - water service which is defined as water located on-premise, available, and free from contamination \citep{WHOdrinkingwater}.   If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water. Another recognition is that many people spend a lot of time collecting water where their access to drinking water
supplies located on-premises are not common, about 26.3 percent of the world's population
\citep{cassivi2018access}. Where this raises concerns about inequality problems related to the task and time variation 
to collect water can be inferred in the urban and rural areas \citep{cassivi2018access}. Therefore, there needs to be a
great motivation to bring improved drinking water sources to all people to improve their quality of life.

@@

Globally, access to improved drinking water sources and sanitation is increasing. From 1990 to 2012, utilization of
improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the MDG categorization,
they consider improved drinking water sources as following: public tap, borehole, protected spring, rainwater
collection, bottled water source, and piped water \citep{bartram2014globala}. However, these variables do not truly
explain if the households have speedy access to retrieve the water. Therefore, we will investigate further on what
variables should be included to enhance the accessibility of improved drinking water.

@@

\section{Relevant work}
Organized by chronology and the Categorical.

\citet{jagals2006does} highlights in his case study in rural area in South Africa discovered providing small communities, using untreated river water as their water source, with good quality water with a piped distribution system and increase of accessibility by installing communal taps did not fall within parameters of safe water. Their research gives a potential understanding even making the drinking water accessible might not improve the quality of the water itself.

\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting impact
on MDG progress. They altered the current Joint monitoring programme (JMP) estimation by
including for microbial water quality and sanitary risk utilizing national water quality data. \citet{onda2012global}
used principal components analysis to analyze the national environment and development indicators and then created
models. Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it
is done manually. How the modes of transportation can affect one's health.

@@

\citet{onda2014country} discusses their new development of typology of country clusters pertain to water and sanitation
sector based on congruence through multiple related indicators. \citet{onda2014country} used a hierarchical clustering
method and a gap statistic analysis to cluster 156 countries which has 6.75 billion people into a relational
clusters. \citet{onda2014country} suggests previous geography or income-based country clustering should be improved by
using water and sanitation related indicators. Our research can improvise upon their finding and focus on identifying
which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.

\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year of 2030 that is
affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
Water Supply and Sanitation (JMP) are monitoring progress on drinking water and sanitation since 1990 globally. The
article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
water and good sanitation. Our study can highlight which modes of transportation of safe and affordable drinking water
can develop why one's community may excel in the improvement of safe and affordable drinking water and
sanitation. Drinking, sanitation, and hand washing ladder explains the qualitative determination can indirectly help
what is a good source of transporting safe and drinking water.

@@

\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which
is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water
access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
temporal dynamics should be considered over three interlinked time scales:
\begin{center}
\begin{tabular}{ c c c }
 short-term & medium term & long-term \\

\end{tabular}
\end{center}

>>>>

\section{Introduction}
In 2017, 29\% of the global population (2.2 billion people) did not use a safely managed drinking-water service, which is defined as water located on-premises, available, and free from contamination \citep{WHOdrinkingwater}.   If you meet about 100 people, you can imagine that 29 of the people you meet do not have access to improved drinking water. Where this raises concerns about inequality problems related to the task and time variation 
to collect water can be inferred in the urban and rural areas \citep{cassivi2018access}. Therefore, there needs to be a
great motivation to bring improved drinking water sources to all people to improve their quality of life.

@@

Globally, access to improved drinking water sources and sanitation is increasing. From 1990 to 2012, the utilization of
improved drinking water sources has increased globally from 76 percent to 89 percent, and utilization of improved
sanitation has increased from 45 percent to 64 percent \citep{fuller2016tracking}. According to the MDG categorization,
they consider improved drinking water sources as the following: public tap, borehole, protected spring, rainwater
collection, bottled water source, and piped water \citep{bartram2014globala}. However, these variables do not truly
explain if the households have speedy access to retrieve the water. In this paper, we investigate further on what
variables should be included to enhance the accessibility of improved drinking water.

@@

\section{Relevant work}

\citet{jagals2006does} highlights in a case study in rural area in South Africa discovered providing small
communities, using untreated river water as their water source, with good quality water with a piped distribution system
and increase of accessibility by installing communal taps did not fall within parameters of safe water. Their research
gives a potential understanding even making the drinking water accessible might not improve the quality of the water
itself.

\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting
impact on MDG progress. They altered the current Joint Monitoring Programme (JMP) estimation by accounting for microbial
water quality and sanitary risk utilizing national water quality data. 
%\citet{onda2012global} used principal components
%analysis to analyze the national environment and development indicators and then created models. 
Overall, their analysis highlights the potential fatigue illness caused by people transporting the water if it is done manually. How the
modes of transportation can affect one's health.

@@

\citet{onda2014country} proposed a country typology based on water and sanitation indicators. They applied hierarchical clustering to 156 countries, accounting for 6.75 billion people. This work improves on previous geography or income-based country clustering. 
%Their research can improvise upon their finding and focus on identifying which water accessibility variables may improve water accessibility for people who are seeking clean drinking water.

\citet{WHOdrinkingwater} article gives an ambitious plan to reach universal drinking water by the year of 2030 that is
affordable and safe followed by the Sustainable Development Goal six goals. The WHO/UNICEF Joint Monitoring Program for
Water Supply and Sanitation (JMP) have been globallhy monitoring progress on drinking water and sanitation since 1990. The
article outlines different plans with consistent traits to determine what is deemed to be safe and affordable drinking
water and good sanitation. Our study can highlight which modes of transportation of safe and affordable drinking water
can develop why one's community may excel in the improvement of safe and affordable drinking water and
sanitation. Drinking, sanitation, and hand washing ladder explains the qualitative determination can indirectly help
what is a good source of transporting safe and drinking water.

@@

\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which
is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water
access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
temporal dynamics should be considered over three interlinked time scales: short, medium and long term.
